---
title: "Summary of exploratory tasks"
author: 
- Ella Yang
- Jimmy Wu
- Alex Morifusa
- Naira Younas
date: today
---
```{r, message = FALSE, warning = FALSE}
library(readr)
library(dplyr)
library(here)
library(knitr)
library(kableExtra)
```


### HTML scraping

```{r, message=FALSE, warning=FALSE}
here::i_am("writeups/tasks-summary.qmd")
res <- readRDS(here("results", "task1-logpcr-summaries.rds"))

res %>%
arrange(metric, desc(mean)) %>%
mutate(mean = round(mean, 4), std_err = round(std_err, 4)) %>%
kbl(
caption = "Logistic PCR (5-fold CV, max_tokens=30000, PCs=300, grid=20)",
align = "lccc"
) %>%
kable_classic(full_width = FALSE)
```


#### 1) Are binary class predictions improved using logistic PCR?
Using logistic PCR with TF-IDF + PCA gives solid performance on the binary task ($\approx$ 0.76 accuracy and 0.82-0.83 ROC-AUC). According to our study, logistic PCR provides a strong baseline for text classification and is the method we use for the comparison below.

#### 2) Does including header content improve predictions?
Yes. Adding headers increases Accuracy by +0.0038 (from 0.7611 to 0.7573) and ROC-AUC by +0.0074 (from 0.8279 to 0.8205).

These gains are small as they are around $\approx$ 0.4-0.7 percentage points but are consistent across both metrics. Given the reported standard errors of around 0.009-0.011, the improvement is fair but not as meaningful. However, the ROC-AUC increase is comparable to the SE and likely meaningful in practice, so we recommend including header text in the pipeline.

For the binary claims classifier, logistic PCR achieves around 76\% CV accuracy. Including headers alongside paragraphs yields a small, consistent improvement in both accuracy and ROC-AUC and should be considered in subsequent modeling.


Does including header content improve predictions? 

### Bigrams

Do bigrams capture additional information relevant to the classification of interest? Answer the question, **briefly** describe what analysis you conducted to arrive at your answer, and provide quantitative evidence supporting your answer.

### Neural net

Summarize the neural network model you trained by desribing:

-   architecture

-   optimization and loss

-   training epochs

-   predictive accuracy
